Edge sign prediction in general graphs and Correlation Clustering

### 2017-08-25 cc/cc_state_of_art.tex
reorganize CC state of the art

As we saw, solving \pcc{} on a training set would give us a principled way of predicting edge signs.
In this section, we thus look at existing methods for \pcc{}. We start with exact methods and fixed
parameter algorithms on complete graphs, noting they implicitly assume our learning bias. However,
in practice, these methods do not scale well, especially if their assumptions are not met. Indeed,
we next present hardness results about \pcc{}, namely that it is \NPh{} to approximate with some
constant factor greater than one, and then describe some proposed approximations. Whereas this
sounds like bad news for us, we show that heuristics can solve large instances with satisfactory
results. We conclude by further highlighting the learning aspect of \pcc{}, describing approaches in
the active and online settings.

@@

@@

In the cluster editing problem, one is given a general unsigned input graph $H=(V,E)$ and wants to
find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
union of cliques. This is equivalent to \pcc{} on complete graphs. To see why, let $G=(V,(E^+,E^-))$
be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
are deleted from $H$.  To the best of our knowledge, the problem was first introduced under this
name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
whereas we shall only give important and more recent pointers here.} They show the problem is
\NPc{}, even if the number of clusters $k\geq2$ is set beforehand\footnote{The reduction is from
3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
$0.878$-approximation in the weighted $k=2$ case using the standard Goemans-Williamson SDP
relaxation.

If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is,
the number of disagreements), then it can be solved in polynomial time in the size of the input
graph (but not in $d$). The best known approximation so far is
$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triples (\ie unbalanced triangles
in the sign language) and branch by either deleting or merging one positive edge. If we additionally
look for exactly $k$ clusters (\ie \mind$[k]$), there is a fixed parameters algorithm running in
$O(2^{O(\sqrt{kd})}+n+m)$~\autocite{Fomin2014}. If at most $a$ edges can be added and at most $b$
edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
the problem can be solved in polynomial time~\autocite{Abu-Khzam2015}. Finally, for planar graph,
there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}.

See \url{http://fpt.wikidot.com/fpt-races} and
\url{http://fpt.wikidot.com/complexity-status-of-edge-modification-problems} for parameterized
papers on \mmc{} and \textsc{Cluster Editing}.

@@

Because of the complexity of the \pcc{} we presented earlier, one has to
rely on approximations to solve large instances of the problem. However, we can imagine offline
signed social networks with only few nodes, in which case it is reasonable to expect finding the
optimal clustering. Furthermore, this can also be useful to evaluate in practice the quality of
heuristic methods, albeit in non asymptotic setting. The most straightforward idea is to use the
formulation of the LP \eqref{eq:mindLP} with the additional constraints that all variable are
binary.

Assign a binary variable $x_{uv}$ to each pair
of nodes (so that $x_{uv}=x_{vu}$). For a given clustering \cluster{}, let $x_{uv} = 0$ if $u$ and
$v$ are in the same cluster and $x_{uv}=1$ is $u$ and $v$ are in different clusters. Noting that
$1-x_{uv}$ is $1$ if the edge $(u,v)$ is within a cluster and $0$ otherwise, the weighted
number of disagreements is then $w(\cluster{}) = \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) +
\sum_{(u,v)\in E^+} w_{uv}x_{uv}$. By construction, if edges $(u,v)$ and $(v,w)$ are within the same
cluster, then $(v, w)$ is also within that cluster. In terms of $x$ variable, we have that $x_{uv}=0
\wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus require that all
variables are either $0$ or $1$ and respect the triangle inequality. We can thus relax the problem into
\begin{align}
   \label{eq:mindLP}
   \text{minimize } & \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) + \sum_{(u,v)\in E^+} w_{uv}x_{uv} \\
   \text{subject to}& \quad x_{uw} \leq x_{uv} + x_{vw} \nonumber\\
   \phantom{subject to}& \quad 0 \leq x_{uv} \leq 1  \nonumber \\
   \phantom{subject to}& \quad x_{uv} = x_{vu}  \nonumber
\end{align}

\Textcite{ExactMIP13} solve this binary integer program on random instances and show that
depending on the negative edge density, the FICO Xpress solver starts to be unable to finish within
a one-hour time limit when $n \geq 40$.
\Textcite{Aref2016} describe four linear and quadratic  binary integer problems
to model \pcc{} with two clusters along with some preprocessing optimizations. With the Gurobi
software, they solve instances with up to $3200$ edges in less than a second. They also sketch
extensions to weighted graphs and more than two clusters.
In a different direction, \textcite{Berg2015}, encode the linear and quadratic integer formulations
into weighted MaxSAT instances and use the state of the art solver
MaxHS\footnote{\href{http://www.maxhs.org/}{http://maxhs.org}}~\autocite{SATSolver13} to get the
exact solution on instances with at most $1000$ nodes in less than a few hours.

After defining the matrix $A$ by $A_{uv} = w_{uv}^+ - w_{uv}^-$, \textcite{LowRank16} show that the
\maxa{} problem can be written up to constant factor by defining one vector $x_u$ per node as:
\begin{align}
   \label{eq:cc_lowrank}
   \text{maximize } & \sum_{u<v} A_{uv}x_u^Tx_v \\
   \text{subject to}& \quad x_{uw} \in \{e_1,\ldots, e_n\} \nonumber
\end{align}
where $e_i$ are the canonical vectors of $\Rbb{}^n$. They show that when the matrix $A$ is positive
semidefinite of rank $k$, the \pcc{} problem can be solved exactly in $O(n^{k^2})$ time. More
practically, they also give an algorithm that closely approximate the objective of
\eqref{eq:cc_lowrank} in $O(nk)$ time.
On complete graphs, we may also turn on fixed-parameter tractable algorithms, where the parameter
can be the number of disagreements, the number of clusters or the maximum number of disagreements
per nodes, as we will now see.

exact describe the LP here
fpa

basically the problem is tractable when the number of disagreements is small, agreeing with our
bias. Likewise, exact methods works well when small rank, smart polytope in LP

@@

Although the same problem was studied earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
coined the term \pcc{} and were the first to study this problem complexity. Namely, for complete,
unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
give a $17429$-approximation of \mind{} and a PTAS%
\bottomfootnote{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in time polynomial in $n$, a
solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
respect to $x$. Note that the time is not necessarily polynomial in $\epsilon$, so that a running
time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
that, for any $\epsilon \in [0,1]$, runs in
$O(n^2e^{O(\frac{1}{\epsilon^{10}}\log\frac{1}{\epsilon})})$ and returns with probability
$1-\frac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
solution of \maxa{}.

\begin{table}[bt]
   \centering
   \small
   \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
   \begin{tabulary}{187mm}{lCCCC}
      \toprule
               & \multicolumn{2}{c}{\mind{}}   & \multicolumn{2}{c}{\maxa{}}                                   \\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      graph    & weighted                      & unweighted                                                     & weighted                                                     & unweighted                   \\
      \midrule
      Complete & \APXh{}~\autocite{Charikar2003} & \NPc{}~\autocite{Bansal2002}, \APXh{}~\autocite{Charikar2003}  &                                                              & \NPc{}~\autocite{Bansal2002} \\
      General  & \APXh{}~\autocites{Charikar2003}{Demaine2003} &  \APXh{}~\autocites{Charikar2003}{Emanuel2003}   & \multicolumn{2}{c}{\APXh{}~\autocite[Thm. 9]{Charikar2003}} \\
      \bottomrule
   \end{tabulary}
\end{table}

The next year, several authors independently strengthened these results and extended them to
weighted and general graphs, as summarized in \autoref{tab:cc_cpx}. In the most complete paper,
\textcite{Charikar2003} show that on complete graphs, minimizing disagreements is \APXh{},
\enquote{that is, is NP-hard to approximate within some constant factor greater than one} (it would
be nice to provide some intuition why it's more difficult to minimize disagreements but according to
the authors themselves, their reduction is \enquote{somewhat intricate}). They prove the same result
on general graphs, for \mind{} by using a reduction from the multicut problem \autocite[Theorem
8]{Charikar2003}\footnote{This reduction was actually turned into an equivalence, as we describe
in more details in \autoref{ssub:cc_under_stability_assumption}.}

We first give more details on the equivalence between \mmc{} and \pcc{},
following~\autocite{Demaine2006} but omitting the full proofs to avoid redundancy. Recall that given
a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
polynomial reduction from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
are simulated in the following way: include every source and target in a clique of $n$ nodes and
connect that clique to the corresponding source or target by $n$ negative edges.

We next present the polynomial reduction from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
4.4]{Demaine2006})\Todo{give the intuition of the proof} that it takes a linear time to construct a
multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.

@@

which asks for the minimum weight set of edges whose removal in $G$ disconnect the
$k$ pairs $(s_i, t_i)$ and for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem
9]{Charikar2003}. The multicut reduction yields a $O(\log n)$ approximation bound, which is achieved
by rounding the canonical Linear Program~\eqref{eq:mindLP}.
Once the LP \eqref{eq:mindLP} is solved, we interpret $x_{uv}$ as a distance: the larger it is and
the more we want $u$ and $v$ to be in different clusters. We can then use the \regionGrow{}
method~\autocite{RegionGrowing93}. Namely, we pick a ball center $u$ \uar{} with radius \shalf{}: if
the average distance of the nodes in the ball to $u$ is less than $\nicefrac{1}{4}$, the ball forms a
cluster, otherwise $\{u\}$ forms a singleton cluster. We then remove the corresponding nodes from
the graph and repeat until all nodes are clustered. As we shall see, this method has been re used
with variations on the LP or on the two thresholds \shalf{} and $\nicefrac{1}{4}$.

@@

In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
\mind{}, by rounding the same LP and using its solution in randomized algorithm. We will not
describe it in detail since a similar idea was used by \textcite{CCPivotConf05} with a better
approximation. To explain it, we first describe their randomized combinatorial algorithm \ccpivot{},
which gives a $3$-approximation of \mind{} on complete unweighted graphs (it has later been
derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
iteration, we pick a node $u$ \uar{} (called the pivot) and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the  LP~\eqref{eq:mindLP} to obtain different approximation factor depending on the
constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
in the context of consensus clustering.} Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, this
tweaking provide a $2.5$-approximation. Note that unweighted graphs naturally fit into that case, as
each edge is either labeled $+$ or $-$. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the LP~\eqref{eq:mindLP} with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

\begin{table}
   \begin{tabulary}{187mm}{llcLL}
      \toprule
                                &            & $k$   & \mind{}                                                                & \maxa{}                                                     \\
      \midrule
      \multirow{2}{*}{Complete} & unweighted &       & $2.06$ \autocite{Chawla2014}                                           & PTAS from \textcite{Bansal2002}  \\
      \cmidrule(r){4-4}
                                & weighted   &       & $1.5$ (with triangular inequality) \autocite{Chawla2014}               &  and by setting $k=\Omega(1/\epsilon)$ \autocite{Giotis2006}  \\
      \midrule
   \multirow{3}{*}{General}     & unweighted &       & \multicolumn{2}{c}{No specific results for the unweighted case}                                                                       \\
      \cmidrule(r){4-5}
                                & weighted   & $k=2$ & $O(\sqrt{\log n})$ \autocite{Giotis2006}                               & $0.884$ \autocite{Mitra2009}                                  \\
      \cmidrule(r){4-5}
                                & weighted   &       & $O(\log n)$ \autocite{Charikar2003}, optimal under the UCG conjecture  & $0.7666$ \autocite{Swamy2004}                                 \\
      \bottomrule
   \end{tabulary}
   \caption{Best results on various problem.\label{tab:cc_approx}}
\end{table}

This concludes the presentation of the known approximation results on \pcc{}, that we summarize in
\autoref{tab:cc_approx}.

\paragraph{Distributed setting}

>>>>

As we saw, solving \pcc{} on a training set would give us a principled way of predicting edge signs.
In this section, we thus look at existing methods for \pcc{}. We start with exact methods and fixed
parameter algorithms on complete graphs, noting they implicitly assume our learning bias. However,
in practice, these methods do not scale well, especially if their assumptions are not met. Indeed,
we next present hardness results about \pcc{}, namely that it is \NPh{} to approximate with some
constant factor greater than one, and then describe some proposed approximations. Whereas this
sounds like bad news for us, we show that heuristics can solve large instances with satisfactory
results. We conclude by further highlighting the learning aspect of \pcc{} through approaches in
the active and online settings.

@@

Because of the complexity of the \pcc{}, one has to rely on approximations to solve large instances
of the problem. However, we can imagine offline signed social networks with only few nodes, in which
case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
to evaluate in practice the quality of heuristic methods, albeit in non asymptotic setting. The
methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
fixed parameter algorithms we describe now only handle complete graphs, while the next set of
methods also work on general graphs.

@@

Let us first describe the cluster editing problem, where one is given a general unsigned input graph
$H=(V,E)$ and wants to
find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
union of cliques. It is possible to navigate between an instance $H$ of cluster editing and an
instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
are deleted from $H$. To the best of our knowledge, the problem was first introduced under this
name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
whereas we shall only give important and more recent pointers here.} They show the problem is
\NPc{}, even if the number of clusters $K\geq2$ is set beforehand\footnote{The reduction is from
3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
$0.878$-approximation in the weighted $K=2$ case using the celebrated Goemans-Williamson SDP
relaxation~\autocite{MaxCutSDP95}.

If we parameterize the problem by the number $d$ of edges that need to be added or deleted (that is,
the number of disagreements), then it can be solved in polynomial time in the size of the input
graph (but not in $d$). The best known approximation so far is
$O(1.62^d+m+n)$~\autocite{GoldenCE12}, which search for conflict triples (\ie unbalanced triangles
in the sign language) and branch by either deleting or merging one positive edge. If we additionally
look for exactly $K$ clusters (\ie \mind$[K]$), there is a fixed parameters algorithm running in
$O(2^{O(\sqrt{kd})}+n+m)$~\autocite{Fomin2014}. If at most $a$ edges can be added and at most $b$
edges can be deleted at each node, and if the minimum size of a cluster is at least $2(a+b)$, then
the problem can be solved in polynomial time~\autocite{Abu-Khzam2015} by using various reduction
rules. Finally, for planar graph,
there is a PTAS running in $O(n2^{\epsilon^{-1}\log(\epsilon^{-1})})$ obtained by dividing the graph
into independent components of bounded treewidth~\autocite{PlanarCEPTAS17}. In other words, if we
are given a complete signed graph whose positive subgraph is planar, then we can minimize the number
of disagreements arbitrarily close to the optimal, provided we have enough time to do so.

% See \url{http://fpt.wikidot.com/fpt-races} and
% \url{http://fpt.wikidot.com/complexity-status-of-edge-modification-problems} for parameterized
% papers on \mmc{} and \textsc{Cluster Editing}.

@@

On general graphs, a natural idea is to formulate the clustering problem as an integer problem and
solve it optimally in polynomial time.
Assign a binary variable $x_{uv}$ to each pair
of nodes (so that $x_{uv}=x_{vu}$). For a given clustering \cluster{}, let $x_{uv} = 0$ if $u$ and
$v$ are in the same cluster and $x_{uv}=1$ is $u$ and $v$ are in different clusters. Noting that
$1-x_{uv}$ is $1$ if the edge $(u,v)$ is within a cluster and $0$ otherwise, the weighted
number of disagreements is then $w(\cluster{}) = \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) +
\sum_{(u,v)\in E^+} w_{uv}x_{uv}$. By construction, if edges $(u,v)$ and $(v,w)$ are within the same
cluster, then $(v, w)$ is also within that cluster. In terms of $x$ variable, we have that $x_{uv}=0
\wedge x_{vw}=0 \implies x_{uw} = 0$. For $x$ to be a valid cluster assignment, we thus require that all
variables are either $0$ or $1$ and respect the triangle inequality.
\begin{align}
   \label{eq:mindIP}
   \text{minimize } & \sum_{(u,v)\in E^-} w_{uv}(1-x_{uv}) + \sum_{(u,v)\in E^+} w_{uv}x_{uv} \\
   \text{subject to}& \quad x_{uw} \leq x_{uv} + x_{vw} \nonumber\\
   \phantom{subject to}& \quad \leq x_{uv} \in \{0, 1\} \label{eq:mindInteger} \\
   \phantom{subject to}& \quad x_{uv} = x_{vu}  \nonumber
\end{align}
\Textcite{ExactMIP13} solve this binary integer program on random instances and show that
depending on the negative edge density, the FICO Xpress solver starts to be unable to finish within
a one-hour time limit when $n \geq 40$.
\Textcite{Aref2016} describe four linear and quadratic  binary integer problems
to model \pcc{} with two clusters along with some preprocessing optimizations. With the Gurobi
software, they solve instances with up to $3200$ edges in less than a second. They also sketch
extensions to weighted graphs and more than two clusters.
In a different direction, \textcite{Berg2015}, encode the linear and quadratic integer formulations
into weighted MaxSAT instances and use the state of the art solver
MaxHS\footnote{\href{http://www.maxhs.org/}{http://maxhs.org}}~\autocite{SATSolver13} to get the
exact solution on instances with at most $1000$ nodes in less than a few hours.

There is also a fixed parameter algorithm to solve \mind{} for general unweighted graphs. Indeed, we
shall see later than the problem is equivalent to \mmc{}, which in this case asks for the minimal
number $d$ of edges to remove in order to separate every pair of nodes in $S=\{(s_i,t_i)\}_{i=1}^k
\in V^2$. It turns out that \mmc{} can be solved exactly in time $2^{O(d^3)}\cdot
n^{O(1)}$~\autocite{FPAMulticut14}.

After defining the matrix $A$ by $A_{uv} = w_{uv}^+ - w_{uv}^-$, \textcite{LowRank16} show that the
\maxa{} problem can be written up to constant factor by defining one vector $x_u$ per node as:
\begin{align}
   \label{eq:cc_lowrank}
   \text{maximize } & \sum_{u<v} A_{uv}x_u^Tx_v \\
   \text{subject to}& \quad x_{uw} \in \{e_1,\ldots, e_n\} \nonumber
\end{align}
where $e_i$ are the canonical vectors of $\Rbb{}^n$. They show that when the matrix $A$ is positive
semidefinite of rank $k$, the \pcc{} problem can be solved exactly in $O(n^{k^2})$ time. More
practically, they also give an algorithm that closely approximate the objective of
\eqref{eq:cc_lowrank} in $O(nk)$ time.

Reflecting on these methods, we can conclude that \pcc{} is solvable in (near) polynomial time
when the problem instance follow our learning bias. Indeed, fixed parameter algorithms are efficient
when the number of disagreements is low, which corresponds to a quasi consistent clustering.
Likewise, the small rank assumption of the adjacency matrix is equivalent to assuming a consistent
clustering~\autocite[Theorem 13]{LowRankCompletion14}. Even when solving integer programs, one can
use the bias to add inequalities and thus tighten the polytope of admissible solutions. On the other
hand, we will now see that in the worst case, for example if an instance does not obey our bias,
then there is little hope to obtain an optimal solution.

@@

Although the same problem was considered earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
coin the term \pcc{} and are the first to study this problem complexity. Namely, for complete,
unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
give a $17429$-approximation of \mind{} and a PTAS%
\marginpars{An algorithm $\mathcal{A}$ is a \emph{polynomial-time approximation scheme (PTAS)} for a
minimization (respectively maximization) problem $\mathcal{P}$  in NP if given any $\epsilon>0$ and
any instance $x$ of $\mathcal{P}$ of size $n$, $\mathcal{A}$ produces, in time polynomial in $n$, a
solution that is within a factor $1+\epsilon$ (respectively $1-\epsilon$) of being optimal with
respect to $x$. Note that the time is not necessarily polynomial in $\epsilon$, so that a running
time of $O(n^{\frac{1}{\epsilon}})$ would qualify~\autocite[Definition 3.10]{CpxBook99}.}
that, for any $\epsilon \in [0,1]$, runs in
$O(n^2e^{O(\epsilon^{-10}\log \epsilon^{-1})})$ and returns with probability
$1-\nicefrac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
solution of \maxa{}.

The next year, several authors independently strengthened these results and extended them to
weighted and general graphs, as summarized in \autoref{tab:cc_cpx}. In the most complete paper,
\textcite{Charikar2003} show that on complete graphs, in sharp contrast with \maxa{} that admits a
PTAS, \mind{} is \APXh{}, \enquote{that is, is NP-hard to approximate within some constant factor
greater than one}. Unfortunately, it is not easy to give intuition why, for this follows from a
\enquote{somewhat intricate reduction from max 2-colorable subgraph problem on bounded degree
3-uniform hypergraphs}. On general graphs though, both versions of the optimization problem are
\APXh{}. It is showed for \maxa{} by a reduction from MAX 3SAT~\autocite[Theorem 9]{Charikar2003}
and for \mind{} by using a reduction from the \mmc{} problem~\autocite[Theorem 8]{Charikar2003}. In
the case of \mind{}, \textcite{Emanuel2003} show a reduction in the other direction. Because \mmc{}
is an interesting graph cut problem, and its complexity has been well studied, we here give more
details on its equivalence with \pcc{}.

For that we follow~\autocite{Demaine2006} but omit the full proofs to avoid redundancy. Recall that
given a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
polynomial transformation from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
are simulated in the following way: include every source and target in a clique of $n$ nodes and
connect that clique to the corresponding source or target by $n$ negative edges.

\begin{table}[b]
   \centering
   \small
   \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
   \begin{tabulary}{187mm}{lCCCC}
      \toprule
               & \multicolumn{2}{c}{\mind{}}   & \multicolumn{2}{c}{\maxa{}}                                   \\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      graph    & weighted                      & unweighted                                                     & weighted                                                     & unweighted                   \\
      \midrule
      Complete & \APXh{}~\autocite{Charikar2003} & \NPc{}~\autocite{Bansal2002}, \APXh{}~\autocite{Charikar2003}  &                                                              & \NPc{}~\autocite{Bansal2002} \\
      General  & \APXh{}~\autocites{Charikar2003}{Demaine2003} &  \APXh{}~\autocites{Charikar2003}{Emanuel2003}   & \multicolumn{2}{c}{\APXh{}~\autocite[Thm. 9]{Charikar2003}} \\
      \bottomrule
   \end{tabulary}
\end{table}

We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
4.4]{Demaine2006})\Todo{give the intuition of the proof} that it takes a linear time to construct a
multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.

@@

If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
within a factor $c$~\autocite{}. On general graphs, the best known approximation factor in $O(\log
k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
$O(\log|E^-|)=O(\log n^2)=O(\log n)$. This approximation ratio is achieved by first solving the
relaxation of the integer problem \eqref{eq:mindIP}, where the constraint \eqref{eq:mindInteger} is
replaced by $x_{uv}\in [0,1]$. From now on, we refer to that relaxation as the \emph{canonical
\mind{} LP}. Once this canonical \mind{} LP is solved, we interpret $x_{uv}$ as a distance: the
larger it is and the more we want $u$ and $v$ to be in different clusters. We can then use the
\regionGrow{} method~\autocite{RegionGrowing93}. The idea is to pick a random center $u$ and to add
to $u$'s cluster all the nodes at distance less than $r$ from $u$ before removing that cluster and
repeating the process. $r<\shalf$ is chosen adaptively such that the weight of positive edges
leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
inside the cluster.\Todo{The full proof can be found in pages 7--9 of
\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
doesn't have much to do with machine learning.}

@@

In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
\mind{}, by rounding the same LP and using a simpler version of \regionGrow{} with fixed radius.
Namely, we pick a ball center $u$ \uar{} with radius \shalf{}: if the average distance of the nodes
in the ball to $u$ is less than $\nicefrac{1}{4}$, the ball forms a cluster, otherwise $\{u\}$ forms
a singleton cluster. We then remove the corresponding nodes from the graph and repeat until all
nodes are clustered.

Not long after, \textcite{CCPivotConf05} came up with a better approximation. To explain it, we
first describe their randomized combinatorial algorithm \ccpivot{}, which gives a $3$-approximation
of \mind{} on complete unweighted graphs (and was later derandomized while preserving its
approximation guarantee~\autocite{derandomCCPivot08}). At each iteration, we pick a node $u$ \uar{}
(called the pivot) and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the canonical LP to obtain different approximation factor depending on the
constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
in the context of consensus clustering.} Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, this
tweaking provide a $2.5$-approximation. Note that unweighted graphs naturally fit into that case, as
each edge is either labeled $+$ or $-$. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the canonical LP with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

\begin{table}
   \caption{Best current results on \pcc{} problems. The \enquote{easiest} setting is \maxa{} on
      complete graphs, for it admits PTASs. All others cases are \APXh{}. However, we see that on
      the diagonal (that is \mind{} on complete graphs and \maxa{} on general graphs), there exists
      constant factor approximations. This is not the case for the most \enquote{difficult} problem,
      \mind{} on general graphs. \label{tab:cc_approx}}
   \begin{tabulary}{\textwidth}{lLL}
      \toprule
      $G$      &  \mind{}                                                  & \maxa{}                                 \\
      \midrule
      Complete &  $2.06$ (and $1.5$ if $G$ is weighted and weights obey the triangular inequality) \autocite{Chawla2014}
               & PTAS from \textcite{Bansal2002} and by setting $k=\Omega(1/\epsilon)$ in \autocite{Giotis2006}      \\
      General  &  $O(\log n)$ \autocite{Charikar2003}                      & $0.7666$ \autocite{Swamy2004}           \\
      \bottomrule
   \end{tabulary}
\end{table}

\medskip
